---
title: "MATH 216 Homework 4"
author: "Amanda Hotvedt"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
library(leaflet)
library(dplyr)
library(ggplot2)
library(rgdal)
library(readr)
```



## Admistrative:

Please indicate

* Who you collaborated with: Katherine Hobbs, Kyra Gray
* Roughly how much time you spent on this HW so far: 8 hours
* The URL of the RPubs published URL [here](http://rpubs.com/ahotvedt/HW4).
* What gave you the most trouble: 
* Any comments you have: I believe my chloropleth map deliverable identifies the ideal shipping hub for DHL! However, I ended up defining so many different data sets throughout the process, which may be hard for someone who is examining my work to follow/understand. I commented out a lot of instructions to try to explain my workflow - so hopefully that will make sense to you! 



## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Perform your computations here:

# Load Social Explorer Data
social_explorer_data <- read_csv("~/Desktop/Data Science/Homework/HW-4/R11276893_SL050.csv")

# Calculating the total US population
social_explorer_data <- social_explorer_data %>% 
  rename(total_pop = SE_T001_001) %>% 
  mutate(Geo_FIPS_char = as.character(Geo_FIPS)) %>% 
  mutate(USA_total_pop = sum(total_pop))

# Polygons
us_counties <- us_counties()

# Centroids
centroids <- coordinates(us_counties) %>% 
  tbl_df()

# Tidy
us_counties_tidy <- tidy(us_counties, region="geoid")

# Actual data
us_counties_data <- us_counties@data

# Join data with new centroid vectors
us_counties_data2 <- bind_cols(us_counties_data, centroids)

# Join data and the tidy format together
us_counties_combined <- left_join(us_counties_tidy, us_counties_data2, by=c("id"="geoid"))

# Join Social Explorer and Counties data
counties_all <- left_join(us_counties_combined, social_explorer_data, by=c("id" = "Geo_FIPS_char"))

# Filter out Alaska, Hawaii, Puerto Rico, and Fairfax County
counties_all <- counties_all %>% 
  filter(state_name != "Alaska") %>% 
  filter(state_name != "Hawaii") %>% 
  filter(state_name != "Puerto Rico") %>% 
  rename(county_name = name) %>% 
  filter(county_name != "Fairfax")

# Finding the weighted mean of distance and population then summing them.
# This gives us one latitude and one longitude, which should be the 
# ideal hub for the shipping company.
FINAL <- counties_all %>% 
  distinct(V1, V2, total_pop, group, USA_total_pop) %>% 
  ungroup() %>% 
  mutate(weighted_lat = total_pop*V2/(USA_total_pop)) %>% 
  mutate(weighted_long = total_pop*V1/(USA_total_pop)) %>% 
  mutate(hub_long = sum(weighted_long), hub_lat = sum(weighted_lat))

# These numbers came from the hub_long and hub_lat variable in FINAL.
# Obviously defining them this way (essentially hard-coding) isn't a best
# practice, but this allowed me to preserve my counties_all data while
# using the 'distinct' function in the FINAL data set to calculate the
# ideal shipping hub's latitude and longitude.
hub_lat <- 40.6
hub_long <- -101.3

# Defining a distance variable
counties_all <- counties_all %>% 
  mutate(distance = (((lat - hub_lat)^2) + ((long - hub_long)^2))^(1/2))

```


```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.
ggplot(data=counties_all, aes(x=long, y=lat, group=group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkblue") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_long, y=hub_lat, col="red", size=3) +
  # Correct aspect ratio for map
  coord_map(xlim=c(-130,-65), ylim=c(20,50))

         
```





